Automatically generated NE tagged corpora for English and Hungarian

### 2012-04-28 paper/engNERwiki.tex
A few modifications to the paper.
Final version.

\documentclass[11pt]{article}
\usepackage{acl2012}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{url}
\usepackage[utf8]{inputenc}
\DeclareMathOperator*{\argmax}{arg\,max}
\setlength\titlebox{6.5cm}    % Expanding the titlebox

@@

%While the approach in essence is language-independent, the exact details have to be accomodated to the target language.

@@

As of version 3.7, the DBpedia ontology allows multiple superclasses, making a
directed acyclic
graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}. Since
selecting the right superclass, and hence, CoNLL tag for classes with more
than one parent, cannot be reliably done automatically, the class-to-category
mapping had to be determined manually. The only such class in version 3.7,
\texttt{Library}, can be traced back to both \texttt{Place} and
\texttt{Organisation}; its CoNLL tag is \texttt{LOC}.  Using the mapping thus
created, we compile a list that contains all entities in DBpedia tagged with
the appropriate CoNLL category.

@@

The following sub-sections describe how the method explained above can be
improved to increase precision, sentence coverage and to account for
peculiarities in the CoNLL guidelines.

@@

\begin{description}
\item[Derived words] According to the CoNLL guidelines, words derived from NEs are tagged as \texttt{MISC}. We complied with this rule by tagging each entity whose head is not a noun, as well as when the link's anchor text is not contained in the entity's name, as \texttt{MISC}. The most prominent example for such entities are nationalities, which can be linked to their home country, a \texttt{LOC}; e.g.~\textit{Turkish} to \textit{Turkey}. Our solution assigns the correct tag to these entities.
\item[First word in a sentence] As first words are always capitalized, labeling them is difficult if they are unlinked and not contained in the entity alias set. We base the decision on the POS tag of the first word: if it is \texttt{NNP}, we tag it as \texttt{UNK}; otherwise, \texttt{O}.
\item[Reference cleansing] Page titles and anchor texts may contain more than
  just the entity name. Personal titles are part of the entity name in WP, but
  not in CoNLL, and punctuation marks around the entity may become part of the
  link by mistake. We tag all punctuation marks after the entity name as
  \texttt{O}. To handle personal titles, we extracted a list from the WP page
  \textit{List of titles}, which contains titles in many languages. We
  manually removed all titles that also function as given names, such as
  \textit{Regina}. If a link to a \texttt{PER} or \texttt{UNK} entity, or an
  unlinked entity starts with, or consists solely of a title in the list, we
  tag the words that make up the title as \texttt{O}.
\item[Incidental capitalization] Various non-\texttt{NNP} words in English are
  capitalized: names of months, the pronoun \textit{I}, and non-entity
  acronyms such as \textit{RSVP}. While the latter two types are unlikely to
  appear in WP text, we assembled a list of these words and tag them as
  \texttt{O} unless they are part of the alias set.
\end{description}

@@

Although Hungarian is reckoned to be a less resourced language, and it is not
supperted in NLTK, several high quality language processing tools have been
developed for Hungarian in recent years. For tokenization and sentence
segmentation we used an in-house statistical tool tailored for Hungarian. It
has been trained on the largest manually annotated Hungarian corpus
\cite{Csendes:04}, and it handles the peculiarities of Hungarian ortography,
such as the periods placed after numbers in date expressions. Lemmatization
was performed by HunMorph \cite{Tron:05} and HunDisambig, an in-house
disambiguator to select the right analysis based on the word context.

@@

There is another special case in Hungarian: unlike in English, the number of
compund words is quite large, and NEs can also be subject to compounding. In
this case the common noun following the NE is joined with a hyphen, so they
constitute one token. However, the joint common noun can modify the original
sense of NE, depending on the semantics of the common noun. For example in
the compound \textit{Nobel-díj} [`Nobel Prize'] the common noun changes the
labeling from \texttt{PER} to \texttt{MISC}, while
%\textit{SAP-szoftvercég} ['SAP software company'] the common noun does not
%change the labeling, but
in the case of the compound \textit{WorldCom-botrány} [`WorldCom scandal'] the NE tag changes from \texttt{ORG} to \texttt{O}. The solution to this problem is not obvious, and needs more investigation.

@@

%az eredeti DBpedia osztályokat is megtarthatnánk, hiszen nem mindig célravezető a CoNLL osztályok használata, mivel azok elég tág kategóriák, amikbe olyan dolgok is tartoznak, amik nagyon eltérőek is lehetnek

@@

% TODO: the corpora will be published in the META-SHARE repository.
This research was supported by the OTKA grant no. 82333 and the CESAR project under the ICT Policy Support Programme (grant no. 271022). The authors are grateful to Attila Zséder for his work on Wikipedia parsing and to András Kornai for his help.

>>>>

\documentclass[11pt]{article}
\usepackage[english]{babel}
\usepackage{acl2012}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{url}
\usepackage[utf8]{inputenc}
\DeclareMathOperator*{\argmax}{arg\,max}
\setlength\titlebox{6.5cm}    % Expanding the titlebox
\hyphenation{DBpedia}

@@

%While the approach in essence is language-independent, the exact details have to be accommodated to the target language.

@@

As of version 3.7, the DBpedia ontology allows multiple superclasses, making a
directed acyclic
graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}. Since
selecting the right superclass, and hence, CoNLL tag, for classes with more
than one parent cannot be reliably done automatically, the class-to-category
mapping had to be determined manually. The only such class in version 3.7,
\texttt{Library}, can be traced back to both \texttt{Place} and
\texttt{Organisation}; its CoNLL tag is \texttt{LOC}.  Using the mapping thus
created, we compile a list that contains all entities in DBpedia tagged with
the appropriate CoNLL category.

@@

The following sub-sections describe how the method explained above can be
improved to increase precision, sentence coverage and to account for
peculiarities in the English orthography and the CoNLL guidelines.

@@

\begin{description}
\item[Derived words] According to the CoNLL guidelines, words derived from NEs are tagged as \texttt{MISC}. We complied with this rule by tagging each entity whose head is not a noun, as well as when the link's anchor text is not contained in the entity's name, as \texttt{MISC}. The most prominent example for such entities are nationalities, which can be linked to their home country, a \texttt{LOC}; e.g.~\textit{Turkish} to \textit{Turkey}. Our solution assigns the correct tag to these entities.
\item[First word in a sentence] As first words are always capitalized, labeling them is difficult if they are unlinked and not contained in the entity alias set. We base the decision on the POS tag of the first word: if it is \texttt{NNP}, we tag it as \texttt{UNK}; otherwise, \texttt{O}.
\item[Reference cleansing] Page titles and anchor texts may contain more than
  just the entity name. Personal titles are part of the entity name in WP, but
  not in CoNLL, and punctuation marks around the entity may become part of the
  link by mistake. We tag all punctuation marks after the entity name as
  \texttt{O}.

  To handle personal titles, we extracted a list from the WP page
  \textit{List of titles}, which contains titles in many languages. We
  manually removed all titles that also function as given names, such as
  \textit{Regina}. If a link to a \texttt{PER} or \texttt{UNK} entity, or an
  unlinked entity starts with, or consists solely of a title in the list, we
  tag the words that make up the title as \texttt{O}.
  % TODO 2 paragraphs?
\item[Incidental capitalization] Various non-\texttt{NNP} words in English are
  capitalized: names of months, the pronoun \textit{I}, and non-entity
  acronyms such as \textit{RSVP}. While the latter two types are unlikely to
  appear in WP text, we assembled a list of these words and tag them as
  \texttt{O} unless they are part of the alias set.
\end{description}

@@

Although Hungarian is reckoned to be a less resourced language, and it is not
supported in NLTK, several high quality language processing tools have been
developed for Hungarian in recent years. For tokenization and sentence
segmentation we used an in-house statistical tool tailored for Hungarian. It
has been trained on the largest manually annotated Hungarian corpus
\cite{Csendes:04}, and it handles the peculiarities of Hungarian orthography,
such as the periods placed after numbers in date expressions. Lemmatization
was performed by HunMorph \cite{Tron:05} and HunDisambig, an in-house
disambiguator to select the right analysis based on the word context.

@@

There is another special case in Hungarian: unlike in English, the number of
compound words is quite large, and NEs can also be subject to compounding. In
this case the common noun following the NE is joined with a hyphen, so they
constitute one token. However, the joint common noun can modify the original
sense of NE, depending on the semantics of the common noun. For example in
the compound \textit{Nobel-díj} [`Nobel Prize'] the common noun changes the
labeling from \texttt{PER} to \texttt{MISC}, while
%\textit{SAP-szoftvercég} ['SAP software company'] the common noun does not
%change the labeling, but
in the case of the compound \textit{WorldCom-botrány} [`WorldCom scandal'] the NE tag changes from \texttt{ORG} to \texttt{O}. The solution to this problem is not obvious, and needs more investigation.

@@

@@

This research was supported by OTKA grant no. 82333 and the CESAR project under the ICT Policy Support Programme (grant no. 271022). The authors are grateful to Attila Zséder for his work on Wikipedia parsing and to András Kornai for his insightful comments.
